---
title: "Evolution of Protein-Protein Interfaces"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include = TRUE)
library(dplyr)
library(tidyr)
library(readr)
library(stringr)
library(ggplot2)
library(cowplot)

proteins <- read_csv("./data/nonenzymes/master_data_table_pp.csv")
```

### Whole-complex RSA values correlate with site-specific evolutionary rates
```{r}

ggplot(proteins, aes(x=RSA, y=rate)) +
  geom_point(alpha=0.1) +
  geom_smooth() +
  coord_cartesian(xlim=c(0,1), ylim=c(0,2))

```

### Whole-complex WCN shows strongest mean correlation with evolutionary rate
```{r}

protein_cors <- proteins %>% 
  group_by(pdb) %>%
  summarize(rsa.r2 = cor.test(RSA, rate)$estimate^2,
         rsa_mono.r2 = cor.test(RSA_mono, rate)$estimate^2,
         wcn.r2 = cor.test(wcnSC, rate)$estimate^2,
         wcn_mono.r2 = cor.test(wcnSC_mono, rate)$estimate^2)

ggplot(protein_cors %>% gather(comparison, r2, rsa.r2:wcn_mono.r2),
       aes(x=factor(comparison, 
                    levels=c("rsa.r2", 
                             "rsa_mono.r2", 
                             "wcn.r2",
                             "wcn_mono.r2"),
                    labels=c("RSA (whole complex), rate",
                             "RSA (chain), rate",
                             "WCN (whole complex), rate",
                             "WCN (chain), rate")), y=r2, group=comparison)) + 
  geom_violin() +
  geom_jitter(position = position_jitter(0.2)) +
  stat_summary(fun.y=mean, geom="point", color="red", size = 4) +
  xlab("Comparison") +
  ylab("R-squared") +
  coord_flip()

```

### ...Although none of these differences are significant
```{r}

make_scatter <- function(df, x, y) {
  ggplot(df, aes(x,y)) + 
    geom_point() +
    geom_abline(slope=1, intercept = 0) +
    xlim(0,0.5) +
    ylim(0,0.5) +
    coord_equal()
}

scatter1 <- make_scatter(protein_cors, protein_cors$rsa.r2, protein_cors$rsa_mono.r2) +
  labs(x="RSA (whole complex)", y = "RSA (chain)")

t.test(protein_cors$rsa.r2, protein_cors$rsa_mono.r2)

scatter2 <- make_scatter(protein_cors, protein_cors$wcn.r2, protein_cors$wcn_mono.r2) +
  labs(x="WCN (whole complex)", y = "WCN (chain)")

t.test(protein_cors$wcn.r2, protein_cors$wcn_mono.r2)

scatter3 <- make_scatter(protein_cors, protein_cors$wcn_mono.r2, protein_cors$rsa_mono.r2) +
  labs(x="WCN (chain)", y = "RSA (chain)")

t.test(protein_cors$wcn_mono.r2, protein_cors$rsa_mono.r2)

scatter4 <- make_scatter(protein_cors, protein_cors$wcn.r2, protein_cors$rsa.r2) +
  labs(x="WCN (whole complex)", y = "RSA (whole_complex)")

t.test(protein_cors$wcn.r2, protein_cors$rsa.r2)

plot_grid(scatter1, scatter2, scatter3, scatter4, ncol=2)

```

### Protein-protein interface residues evolve more slowly than non-interface residues
```{r}

proteins <- proteins %>%
  mutate(interface = ifelse(RSA_mono - RSA > 0.01, T, F))

ggplot(proteins, aes(x=interface, y=rate)) + 
  geom_violin() +
  stat_summary(fun.y=mean, geom="point", color="red", size = 3) +
  coord_cartesian(ylim=c(0,1))

t.test(filter(proteins, interface==T)$rate, 
       filter(proteins, interface==F)$rate)

```

### Enzyme data set: protein-protein interface residues evolve more quickly than catalytic residues, but more slowly than other residues

```{r}
enzymes <- read_csv("master_data_table.csv") %>%
  mutate(interface = ifelse(RSA_mono - RSA > 0.01, T, F))

ggplot(enzymes, aes(x=interface, y=rate)) + 
  geom_violin() +
  stat_summary(fun.y=mean, geom="point", color="red", size = 3) +
  coord_cartesian(ylim=c(0,1))

t.test(filter(enzymes, interface==T)$rate, 
       filter(enzymes, interface==F)$rate)

cats <- filter(enzymes, ACTIVE_SITE == T) %>%
  mutate(type="cat") %>%
  select(type, rate)
ints <- filter(enzymes, interface == T) %>%
  mutate(type="interface") %>%
  select(type, rate)
nonenzyme <- filter(proteins, interface == T) %>%
  mutate(type="interface_nonenzyme") %>%
  select(type,rate)

cats_ints <- bind_rows(cats, ints, nonenzyme)

ggplot(cats_ints, aes(x=factor(type, 
                               levels=c("cat", "interface", "interface_nonenzyme"),
                               labels=c("Catalytic",
                                        "Interface (enzymes)",
                                        "Interface (nonenzymes)")), y=rate)) +
  geom_violin() +
  coord_cartesian(ylim=c(0,1)) +
  stat_summary(fun.y=mean, geom="point", color="red", size = 4) +
  xlab("Residue Type")

t.test(cats$rate,ints$rate)

t.test(ints$rate, nonenzyme$rate)

```

### Conclusions

- Catalytic residues are highly conserved
- Protein-protein interface residues evolve more quickly than catalytic residues
- Protein-protein interface residues evolve more slowly than non-interface, non-catalytic residues